---
title: "PUMA Modeling Analysis"
output: html_document
date: "2025-04-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Set up
```{r, include=FALSE}
#load packages
library(dplyr)
library(flextable)
library(ggplot2)
library(collapse)
library(labelled)
library(compareGroups)
library(gee)
library(geepack)
library(pROC)
library(ROCR)
library(plotROC)
library(gridExtra)
library(kableExtra)
library(ggalluvial)
```

# Load Data and Source Functions
```{r}
pathway <- "~/Library/CloudStorage/Box-Box/Puma Analysis/Analysis/Code"
month12 <- readRDS(file.path(pathway, "/RDS/puma_month12.rds"))
puma_all <- readRDS(file.path(pathway, "/RDS/puma_long_clean.rds"))
secondary_long <- readRDS(file.path(pathway, "/RDS/puma_secondary_long.rds"))

output_path <- "~/Library/CloudStorage/Box-Box/Puma Analysis/Analysis/Code/Output"

source("analysis_utilities.R")
```

# Prepare Data for GEE models
```{r}
# Create dataframes for the GEE models section
## Hair
hair_gee_data <- prepare_gee_data(
  puma_all, 
  outcome = "detect_tfv", 
  predictors = c("wilson_adh_scaled_sd", "group_new"),
  additional_vars = c("wilson_adh10", "wilson_adh", "hair3numeric", "hair2_coded",
                      "emp_coded", "age", "edu")
)


## Urine
ut_gee_data <- prepare_gee_data(
  puma_all, 
  outcome = "ut_result", 
  predictors = c("wilson_adh_scaled_sd", "group_new"),
  additional_vars = c("wilson_adh", "wilson_adh10", "emp_coded", "age", "edu")
)

## Continuous hair
continuous_data <- prepare_gee_data(
  puma_all, 
  outcome = "tfvngml", 
  predictors = c("wilson_adh_scaled_sd", "group_new"),
  additional_vars = c("wilson_adh", "wilson_adh10", "ptid", "visit", "detect_tfv")
)

## POC all
poc_all <- prepare_gee_data(
  puma_all, 
  outcome = "ut_result", 
  predictors = c("wilson_adh_scaled_sd", "group_new"),
  additional_vars = c("wilson_adh", "wilson_adh10", "ptid", "visit", "detect_tfv")
)

poc_all <- poc_all %>% filter(group_new == "Intervention")

## SOC all
soc_all <- prepare_gee_data(
  puma_all, 
  outcome = "ut_result", 
  predictors = c("wilson_adh_scaled_sd", "group_new"),
  additional_vars = c("wilson_adh", "wilson_adh10", "ptid", "visit", "detect_tfv")
)

soc_all <- soc_all %>% filter(group_new == "Standard of Care")

```

# Hair -- Binary (0/1)
## Adherence Scaled by Standard Deviation
```{r}
hair_sd_results <- run_gee_model(
  hair_gee_data,
  outcome = "detect_tfv",
  predictors = c("wilson_adh_scaled_sd", "group_new")
)

flextable(hair_sd_results)
```

## Adherence Not-Scaled
```{r}
hair_results <- run_gee_model(
  hair_gee_data,
  outcome = "detect_tfv",
  predictors = c("wilson_adh", "group_new")
)

flextable(hair_results)
```

## Adherence Scaled by Factor of 10
```{r}
hair_10_results <- run_gee_model(
  hair_gee_data,
  outcome = "detect_tfv",
  predictors = c("wilson_adh10", "group_new")
)

flextable(hair_10_results)
```

# Hair -- Continuous
## Adherence Scaled by Standard Deviation
```{r}
# Association between adherence and continuous hair levels
# Linear regression with a GEE model 
# Association between adherence and continuous hair levels
hair_cont_results <- run_gee_continuous_model(
  continuous_data,
  outcome = "tfvngml",
  predictors = c("wilson_adh_scaled_sd", "group_new")
)

flextable(hair_cont_results)

```

# Hair -- Categorical with 3 codes (< 2 doses, 2-3 doses, >= 4 doses)
## Adherence Scaled by Standard Deviation
```{r}
hair_3cat_results <- run_gee_model(
  hair_gee_data,
  outcome = "hair3numeric",
  predictors = c("wilson_adh_scaled_sd", "group_new")
)

flextable(hair_3cat_results)
```

# Hair -- Categorical with 2 codes (High [1; >= 4 doses] vs Low [0; <= 3 doses])
## Adherence Scaled by Standard Deviation
```{r}
hair_2cat_results <- run_gee_model(
  hair_gee_data,
  outcome = "hair2_coded",
  predictors = c("wilson_adh_scaled_sd", "group_new")
)

flextable(hair_2cat_results)
```

# Urine -- Binary (0/1)
## Adherence Scaled by Standard Deviation
```{r}
# Urine model with scaled_sd
ut_sd_results <- run_gee_model(
  ut_gee_data,
  outcome = "ut_result",
  predictors = c("wilson_adh_scaled_sd", "group_new")
)

flextable(ut_sd_results)
```

## Adherence Not-Scaled
```{r}
urine_results <- run_gee_model(
  ut_gee_data,
  outcome = "ut_result",
  predictors = c("wilson_adh", "group_new")
)

flextable(urine_results)
```

## Adherence Scaled by Factor of 10
```{r}
ut_10_results <- run_gee_model(
  ut_gee_data,
  outcome = "ut_result",
  predictors = c("wilson_adh10", "group_new")
)

flextable(ut_10_results)
```

# Multivariate Analysis
## Univariable Analysis 
```{r}
# List of predictor variables selected a priori
predictors <- c("wilson_adh", "wilson_adh_scaled_sd", "padh4", "padh7",
                "ppoz_coded", "marital", "income", "travel_coded", 
                "incsource_coded", "edu", "emp_coded", "mh14", "sex1", 
                "prep_length", "condomless_sex", "sex14", "age")

# List of outcome variables
outcomes <- c("ut_result", "hair2_coded")

# Run univariable analysis
univariable_results <- run_univariable_analysis(
  data = secondary_long,
  predictors = predictors,
  outcomes = outcomes
)

# Get summary of significant predictors
significant_predictors <- summarize_univariable_results(univariable_results)

# Display the full results table
print(univariable_results)
## Associated univariable with ut:
### emp_coded, age

## Associated univariable with detect_tfv
### edu 
```

## Univariable Models with Signficiant Associated Variables
### Urine
```{r}
## emp_coded
emp_results <- run_gee_model(
  secondary_long,
  outcome = "ut_result",
  predictors = c("emp_coded")
)

flextable(emp_results)

## age
age_results <- run_gee_model(
  secondary_long,
  outcome = "ut_result",
  predictors = c("age")
)

flextable(age_results)
```

### Hair
```{r}
## edu
edu_results <- run_gee_model(
  secondary_long,
  outcome = "hair2_coded",
  predictors = c("edu")
)

flextable(edu_results)

## age
age_results_h <- run_gee_model(
  secondary_long,
  outcome = "hair2_coded",
  predictors = c("age")
)

flextable(age_results_h)
```

## Forward Selection with Significant Variables
### Urine
```{r}
forward_selection_vars <- c("edu", "emp_coded", "age", "hair2_coded", "ut_result", 
                            "group_new", "wilson_adh_scaled_sd", "ptid")

multivar_data <- secondary_long[forward_selection_vars]

### Urine Assay
# Remove NAs
multivar_data <-multivar_data[complete.cases(multivar_data),]
ut_multivar <- multivar_data %>% select(-hair2_coded)

## Manual stepwise

# Start with minimum model
min_model_multivar <- geeglm(ut_result ~ group_new,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = ut_multivar)
QIC(min_model_multivar)

# Add in Wilson ADH
wilson_model_multivar <- geeglm(ut_result ~ group_new + wilson_adh_scaled_sd,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = ut_multivar)
QIC(wilson_model_multivar)

# Add in employment status 
emp_model_multivar <- geeglm(ut_result ~ group_new + wilson_adh_scaled_sd + emp_coded,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = ut_multivar)
# QIC is higher thn just Wilson ADH
QIC(emp_model_multivar)

# Remove emp and add in age
age_model_multivar <- geeglm(ut_result ~ group_new + wilson_adh_scaled_sd + age,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = ut_multivar)
# Compare QICs
QIC(wilson_model_multivar)
# Age is lower 
QIC(age_model_multivar)

# Add all in
full_model_multivar <- geeglm(ut_result ~ group_new + emp_coded + age + wilson_adh_scaled_sd,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = ut_multivar)
# Adding in emp has higher QIC
QIC(full_model_multivar)

## Print full model with age
full_m_ut <- run_gee_model(
  secondary_long,
  outcome = "ut_result",
  predictors = c("group_new", "wilson_adh_scaled_sd", "age")
)

flextable(full_m_ut)
```

### Hair Forward Selection
```{r}
### Urine Assay
# Remove NAs
multivar_data <-multivar_data[complete.cases(multivar_data),]
hr_multivar <- multivar_data %>% select(-ut_result)

## Manual stepwise

# Start with minimum model
min_model_hr <- geeglm(hair2_coded ~ group_new,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = hr_multivar)
QIC(min_model_hr)

# Add in Wilson ADH
wilson_model_hr <- geeglm(hair2_coded ~ group_new + wilson_adh_scaled_sd,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = hr_multivar)
QIC(wilson_model_hr)

# Add in edu
edu_model_hr <- geeglm(hair2_coded ~ group_new + wilson_adh_scaled_sd + edu,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = hr_multivar)
# QIC is lower than just Wilson ADH
QIC(edu_model_hr)

# Just edu by itself
edu_model2_hr <- geeglm(hair2_coded ~ group_new + edu,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = hr_multivar)
# QIC is higher than above
QIC(edu_model2_hr)

# Add in age
age_edu_model_hr <- geeglm(hair2_coded ~ group_new + wilson_adh_scaled_sd + edu + age,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = hr_multivar)
# QIC is lower than just Wilson ADH
QIC(age_edu_model_hr)

## Print full model with education status
full_m_hr <- run_gee_model(
  secondary_long,
  outcome = "hair2_coded",
  predictors = c("group_new", "wilson_adh_scaled_sd", "edu")
)

flextable(full_m_hr)
```

## Full Models for Both Results
```{r}
#Urine
model_ut <- run_gee_model(
  secondary_long,
  outcome = "ut_result",
  predictors = c("group_new", "wilson_adh_scaled_sd", "edu", "age")
)

flextable(model_ut)

## Check QIC using geeglm
m_ut_glm <- geeglm(ut_result ~ group_new + edu + age + wilson_adh_scaled_sd,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = secondary_long)
QIC(m_ut_glm)

#Hair
m_hair <- run_gee_model(
  secondary_long,
  outcome = "hair2_coded",
  predictors = c("group_new", "wilson_adh_scaled_sd", "edu", "age")
)

flextable(m_hair)

## Check QIC using geeglm
mhair_glm <- geeglm(hair2_coded ~ group_new + edu + age + wilson_adh_scaled_sd,
                id = ptid,
                family = poisson(link = "log"),
                corstr = "exchangeable",
                data = hr_multivar)
QIC(mhair_glm)
```

# Stratified Analysis
## Hair
```{r}
# Filter data for analysis (identical to original)
hair_gee_data <- puma_all %>%
  filter(complete.cases(detect_tfv, wilson_adh, group_new, ptid)) %>% 
  select(detect_tfv, wilson_adh_scaled_sd, wilson_adh10, 
         group_new, ptid, wilson_adh, mh14, hair3numeric)

# Analyze intervention group
hair_poc_results <- run_group_specific_gee(
  data = hair_gee_data,
  outcome = "detect_tfv",
  predictor = "wilson_adh_scaled_sd",
  group_var = "group_new",
  group_value = "Intervention"
)

# Display results
flextable(hair_poc_results)

# Analyze standard of care group
hair_soc_results <- run_group_specific_gee(
  data = hair_gee_data,
  outcome = "detect_tfv",
  predictor = "wilson_adh_scaled_sd",
  group_var = "group_new",
  group_value = "Standard of Care"
)

# Display results
flextable(hair_soc_results)
```

## Urine 
```{r}
# Filter data for urine analysis (identical to original)
ut_gee_data <- puma_all %>%
  filter(complete.cases(ut_result, wilson_adh, group_new, ptid)) %>% 
  select(ut_result, wilson_adh_scaled_sd, wilson_adh10, 
         group_new, ptid, wilson_adh, mh14)

# Analyze standard of care group
ut_soc_results <- run_group_specific_gee(
  data = ut_gee_data,
  outcome = "ut_result",
  predictor = "wilson_adh_scaled_sd",
  group_var = "group_new",
  group_value = "Standard of Care"
)

# Display results
flextable(ut_soc_results)

# Analyze intervention group
ut_poc_results <- run_group_specific_gee(
  data = ut_gee_data,
  outcome = "ut_result",
  predictor = "wilson_adh_scaled_sd",
  group_var = "group_new",
  group_value = "Intervention"
)

# Display results
flextable(ut_poc_results)
```

## Interaction Tests for Difference between 
```{r}
# Run interaction test for urine
ut_lrt_results <- run_group_interaction_test(
  data = ut_gee_data,
  outcome = "ut_result",
  predictor = "wilson_adh_scaled_sd",
  group_var = "group_new"
)

# Display results
flextable(ut_lrt_results)

```

```{r}

```

